### 2020-03-30 rebuttal.txt
Import initial rebuttal by Enrico
Co-Authored-By: Enrico Tassi <enrico.tassi@inria.fr>

Here the message I propose, it is ~600 words so there is space if you
want to write more:

Dear PC members, first of all we would like to thank the reviewers for
their valuable feedback. We believe we will be able to address all
their requests for the final version, namely clarify the exposition in
sections 4 and 5 (as per review 1 and 2), and add a few extra details
in section 6 (as per review 2).

Specific answer to reviewer 1
-----------------------------

- p4-5, evolution of the hierarchy:
     Could you make clearer the distinction (or absence of) between
     axioms and theorems in your implementation of a theory (problem
     mentioned on line 121). I.e., does it matter or not for the user
     of the hierarchy.

It does not matter: addrN is name he can use in her proofs. What
changed is how it is "implemented" by the hierarchy designer. Line
121 is very much related to the remark at lines 421-425, we shall
hint that in section 2 as well.

- Propositions 5 and 8: there is no proof, but is there anything to
prove?

We agree that 5 is a remark. On the contrary 8 shows M \subset F
by providing an implementation of requires, provides and from.
This proposition identifies *factories* as the central notion of HB,
since mixins are just a special case. Trivia: in an early draft the
name of the language was, indeed, Factory.

- Section 4:
     What are you describing here?
     Is it the code your tool generate?
     Does it correspond (exactly) to examples in Section 2?
     Has anything been omitted?

4.1 show Coq code that is equivalent to the HB commands issued
in Section 2.
HB commands do not synthesize that text, but have the same effect.

Many details were omitted. For example the abbreviations mentioned in
section 3.3 are not there. Another omission is that all "exported"
notions are gathered into an inner module, eg Monoid.Exports, etc.
We believe these details are not very interesting, but we will clarify
the status of the Coq code in section 4.

General answer to all reviewers
-------------------------------
All reviewers acknowledge that HB looks useful. Reviewer 1 flatters
us with "the tool is invaluable" and acknowledges that a contribution
of HB is "providing a way for normal Coq users to write hierarchies
using the packed class discipline". Reviewer 2 finds it "promising".
Reviewer 3 finds the machinery "impressive".

The main critic to HB, by Reviewer 3, is that it looks not only
impressive,
but also intimidating and Reviewer 3 wonders if a Coq user would run
the
risk of depending on HB in his work.

We use the remaining space to comment on that, that we believe is a
very pertinent remark.

In an early prototype of HB we considered generating text (Coq source
code).
We believe this approach could let users take advantage of HB without
depending on it (permanently). We do not provide this feature in the
current code because, honestly, it is not the most intriguing feature
to
implement or to describe in a paper.

We shall consider that feature again in the future, as an extreme
measure,
if the software stack behind HB turns out to severely hinder its
adoption.
We hope Reviewer 1 is right in saying that even regular Coq users could
be interested in using HB, and if the tool turns out to be popular its
maintenance will be assured by its authors and the community of users,
as it is the case for the Coq system or the Mathematical Components
library.