### 2020-03-30 rebuttal.txt
reworked from google docs

Dear PC members, first of all we would like to thank the reviewers for their
valuable feedback. We believe we will be able to address all their requests for
the final version, namely clarify the exposition in sections 4 and 5 (as per
review 1 and 2), and add a few extra details in section 6 (as per review 2).

Answer to reviewer 1
-----------------------------

- p4-5, evolution of the hierarchy:
         Could you make clearer the distinction (or absence of) between
         axioms and theorems in your implementation of a theory (problem
         mentioned on line 121). I.e., does it matter or not for the user
         of the hierarchy.

There are two distinct use cases for a hierarchy:
1. one assumes a structure as
     an hypothesis. In this case, all its axioms are theorems so the distinction
     axiom-theorem does not matter.
2. one instantiates a structure on a concrete
     type in order to access all its theory. Here, however, one must pick an
     axiomatic (i.e. a set of axioms) to show Coq that some type is indeed an
     instance of a structure. If for each structure there is only one axiomatic
     (which is its internal representation) and if we keep changing it, some theorems
     may become axioms and increase the burden on the user (this is what happens with
     addrC in V2, in order to instantiate a ring one has to prove it, while it was
     not required in V1). We fix that by letting the designer provide as many
     axiomatics as she wants, so that the user can pick the most convenient one, in
     particular V3 also provides the axiomatic from V1 as a factory. Behind the
     scenes, hierarchy-builder compiles it to the axiomatic currently used as the
     internal representation (made of mixins).

We plan to integrate a variation of the text above as the transition from
section 2.1 to 2.2.

- Propositions 5 and 8: there is no proof, but is there anything to
Prove?

Albeit trivial, they are both theorems. Our mistake was to mix the proof with
the statement.

Proposition 5) dep represents a transitively closed DAG, in particular
  dep*(dep*(M)) ⊆ dep*(M).
Proof.
dep is transitively closed since records are well typed in the empty context.
dep describes a DAG since Coq does not admit circular definitions.
Qed.

Proposition 8) There is a way to see mixins as factories (M ⊆  F).
Proof.
For each m ∈ M we pose
* requires(m) = dep(m)
* provides(m) = {m}
* from (m, m) = fun T (p : m T p σ ) (x : m T p 1 . . . p n ) => x ∈ B
Qed.

This proposition identifies *factories* as the central notion of HB, since
mixins are just a special case. Trivia: in an early draft the name of the
language was, indeed, Factory.

- Section 4:
         What are you describing here?
         Is it the code your tool generates?
         Does it correspond (exactly) to examples in Section 2?
         Has anything been omitted?

Section 4.1 shows Coq code that is equivalent to the HB commands given in
Section 2. These commands do not synthesize that text, but have the same effect.

Many details were omitted:
* the abbreviations mentioned in section 3.3
* the declarations of canonical structures are postponed to 4.3
* "Exported" notions are gathered into an inner module, eg Monoid.Exports.

We believe these details are not very interesting, but we will clarify the
status of the Coq code in section 4 in the final version.

Comment to all reviewers
-------------------------------

We could not avoid to remark that, by lack of clarity on our side, an important
aspect of HB was not sufficiently put forward. Indeed two reviewers write
seemingly opposite remarks:

* Reviewer 1 : "The first [contribution], surprisingly little advertised, [is]
     providing a way for normal Coq users to write hierarchies using the packed class
     discipline"

* Reviewer 3 :  “I can imagine a few experts using these tools to build
     hierarchies, but I would not expect the average Coq user to invest time into
     learning about them”.

We agree sections 4 and 5 of the paper can look intimidating to a non Coq +
Packed Classes expert, but these details are essentially the ones a user of HB
does not need to understand in order to use the tool. Indeed, the implementation
of HB takes care of these details and avoids most of the common mistakes one can
hardly avoid when writing Packed Classes code by hand. In the final version of
the paper, we intend to clarify this aspect in the introduction.

Finally, It is true that the software stack is large and may discourage the
adoption of HB, as remarked by Reviewer 3. In an early prototype of HB we
considered generating text (Coq source code). We believe this approach could let
users take advantage of HB without depending on it (permanently). We do not
provide this feature in the current code base because, honestly, it is not the
most intriguing feature to implement or to describe in a paper. We shall
consider that feature again in the future, as an extreme measure, if the
software stack behind HB turns out to severely hinder its adoption. However, we
hope Reviewer 1 is right in saying that the tool is invaluable, and if the tool
turns out to be popular its maintenance will be assured by its authors and the
community of users, as it is the case for the Coq system or the Mathematical
Components library.