### 2020-03-30 rebuttal.txt
typos, clarifications, typesetting

Dear PC members, first of all we would like to thank the reviewers for their
valuable feedback. We believe we will be able to address all their requests for
the final version, namely clarify the exposition in sections 4 and 5 (as per
review 1 and 2), and add a few extra details in section 6 (as per review 2).

*Answer to reviewer 1*

- p4-5, evolution of the hierarchy: Could you make clearer the
  distinction (or absence of) between axioms and theorems in your
  implementation of a theory (problem mentioned on line 121). I.e.,
  does it matter or not for the user of the hierarchy.

There are two distinct use cases for a hierarchy:
1. one assumes a structure as an hypothesis. In this case, all its
   axioms are theorems so the distinction axiom-theorem does not
   matter.
2. one instantiates a structure on a concrete type in order to access
   all its theory. Here, however, one must pick an axiomatic (i.e. a
   set of axioms) to show Coq that some type is indeed an instance of
   a structure. If for each structure there is only one axiomatic
   (which is its internal representation) and if we keep changing it
   across versions, some theorems may become axioms and increase the
   burden on the user (this is what happens with addrC, in order to
   instantiate a ring in V2 one has to prove it, while it was not
   required in V1). We fix that by letting the designer provide as
   many axiomatics (i.e. factories) as she wants, so that the user can
   pick the most convenient one, in particular V3 also provides the
   axiomatic from V1. Behind the scenes, hierarchy-builder compiles it
   to the axiomatic currently used as the internal representation
   (made of mixins).

We plan to integrate a variation of the text above as the transition from
section 2.1 to 2.2.

- Propositions 5 and 8: there is no proof, but is there anything to Prove?

Albeit trivial, they are both theorems. Our mistake was to mix the proof with
the statement. Here is the modification we will make:

Proposition 5) dep represents a transitively closed DAG, in particular
  dep*(dep*(M)) ⊆ dep*(M).
Proof.
dep is transitively closed since records are well typed in the empty context.
dep describes a DAG since Coq does not admit circular definitions.
Qed.

Proposition 8) There is a way to see mixins as factories (M ⊆  F).
Proof.
For each m ∈ M we pose
* requires(m) = dep(m)
* provides(m) = {m}
* from (m, m) = fun T (p : m T p σ ) (x : m T p 1 . . . p n ) => x ∈ B
Qed.

The latter proposition identifies *factories* as the central notion of
HB, since mixins are just a special case. Trivia: in an early draft
the name of the language was, indeed, Factory.

- Section 4: What are you describing here?  Is it the code your tool
  generates?  Does it correspond (exactly) to examples in Section 2?
  Has anything been omitted?

Section 4.1 shows Coq code that is almost equivalent to the HB commands given in
Section 2. These commands do not synthesize that text, but have the same effect.

Many details were omitted:
- the abbreviations mentioned in section 3.3
- the declarations of canonical structures are postponed to 4.3
- "Exported" notions are gathered into an inner module, e.g. Monoid.Exports.

We believe these details are not very interesting, and we will clarify the
status of the Coq code in section 4 in the final version.

*Comment to all reviewers*

We could not avoid to remark that, by lack of clarity on our side, an important
aspect of HB was not sufficiently put forward. Indeed two reviewers write
seemingly opposite remarks:
- Reviewer 1 : "The first [contribution], surprisingly little
   advertised, [is] providing a way for normal Coq users to write
   hierarchies using the packed class discipline"
- Reviewer 3 : “I can imagine a few experts using these tools to build
   hierarchies, but I would not expect the average Coq user to invest
   time into learning about them”.

We agree sections 4 and 5 of the paper can look intimidating to a non Coq +
Packed Classes expert, but these details are essentially the ones a user of HB
does not need to understand in order to use the tool. Indeed, the implementation
of HB takes care of these details and avoids most of the common mistakes one can
hardly avoid when writing Packed Classes code by hand. In the final version of
the paper, we intend to clarify this aspect in the introduction.

Finally, It is true that the software stack is large and may discourage the
adoption of HB, as remarked by Reviewer 3. In an early prototype of HB we
considered generating text (Coq source code). We believe this approach could let
users take advantage of HB without depending on it (permanently). We do not
provide this feature in the current code base because, honestly, it is not the
most intriguing feature to implement or to describe in a paper. We shall
consider that feature again in the future, as an extreme measure, if the
software stack behind HB turns out to severely hinder its adoption. However, we
hope Reviewer 1 is right in saying that the tool is invaluable, and if the tool
turns out to be popular its maintenance will be assured by its authors and the
community of users, as it is the case for the Coq system or the Mathematical
Components library.