### 2020-06-19 coqws/easychair.tex
Improving the presentation.

% easychair.tex,v 3.5 2017/03/15

\documentclass{easychair}
%\documentclass[EPiC]{easychair}
%\documentclass[EPiCempty]{easychair}
%\documentclass[debug]{easychair}
%\documentclass[verbose]{easychair}
%\documentclass[notimes]{easychair}
%\documentclass[withtimes]{easychair}
%\documentclass[a4paper]{easychair}
%\documentclass[letterpaper]{easychair}
\usepackage{amsmath, amssymb}

\usepackage{wrapfig}
\usepackage[font=small,labelfont=bf]{caption}

\usepackage{doc}
\usepackage{minted} % pip3 install pygments, pdflatex -shell-escape hb
\newminted[elpicode]{elpi.py:ElpiLexer -x}{linenos=true,fontsize=\footnotesize,escapeinside=\%\%}
\newmintinline[elpi]{elpi.py:ElpiLexer -x}{fontsize=\small}
\newminted[coqcode]{elpi.py:CoqElpiLexer -x}{firstnumber=last,linenos=false,fontsize=\footnotesize,mathescape,escapeinside=\#\#}
\newmintinline[coq]{elpi.py:CoqElpiLexer -x}{fontsize=\small}
\definecolor{dkgreen}{rgb}{0, 0.502, 0}
\bibliographystyle{plainurl}% the mandatory bibstyle

% use this if you have a long article and want to create an index
% \usepackage{makeidx}

% In order to save space or manage large tables or figures in a
% landcape-like text, you can use the rotating and pdflscape
% packages. Uncomment the desired from the below.
%
% \usepackage{rotating}
% \usepackage{pdflscape}

% Some of our commands for this guide.
%
\newcommand{\easychair}{\textsf{easychair}}
\newcommand{\miktex}{MiK{\TeX}}
\newcommand{\texniccenter}{{\TeX}nicCenter}
\newcommand{\makefile}{\texttt{Makefile}}
\newcommand{\latexeditor}{LEd}

%\makeindex

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Hierarchy Builder}

% Authors are joined by \and. Their affiliations are given by \inst, which indexes
% into the list defined using \institute
%

\author{Cyril Cohen\inst{1} \and Kazuhiko Sakaguchi\inst{2} \and Enrico Tassi\inst{3}}

\institute{
 Inria, Universit\'e C\^ote d'Azur, France\\ \email{Cyril.Cohen@inria.fr}
\and
 University of Tsukuba, Japan\\ \email{sakaguchi@logic.cs.tsukuba.ac.jp}
\and
 Inria, Universit\'e C\^ote d'Azur, France\\\ \email{Enrico.Tassi@inria.fr}
}

%  \authorrunning{} has to be set for the shorter version of the authors' names;
% otherwise a warning will be rendered in the running heads. When processed by
% EasyChair, this command is mandatory: a document without \authorrunning
% will be rejected by EasyChair

\authorrunning{C.\,Cohen and K.\,Sakaguchi and E.\,Tassi}

% \titlerunning{} has to be set to either the main title or its shorter
% version for the running heads. When processed by
% EasyChair, this command is mandatory: a document without \titlerunning
% will be rejected by EasyChair
\titlerunning{Hierarchy Builder}
\newcommand{\HB}{\ensuremath{\mathcal{HB}}}

\begin{document}

\maketitle

%\begin{abstract}
%  We present a tool that makes it easy to build hierarchies
%  of algebraic structures in Coq.
%\end{abstract}

% The table of contents below is added for your convenience. Please do not use
% the table of contents if you are preparing your paper for publication in the
% EPiC Series or Kalpa Publications series

%\setcounter{tocdepth}{2}
%{\small
%\tableofcontents}

%\section{To mention}
%
%Processing in EasyChair - number of pages.
%
%Examples of how EasyChair processes papers. Caveats (replacement of EC
%class, errors).

\paragraph{Context and Talk proposal}

Abstract algebra is extremely important in the mathematical vernacular
and libraries of machine checked code are, nowadays, organized around
hierarchies of structures. Unfortunately the language of Type Theory
and the features provided by the Coq system make the construction of a
hierarchy hard even for expert users. The difficulty begins with the
non-orthogonal choices, between storing information as record fields
or parameters, and between using type classes and canonical structures
for inference. To this, one may add the concerns about performance and
about the usability, by a non expert, of the final hierarchy.

\HB{} gives the library designer a language to describe the building
blocks of algebraic structures and to assemble them into a
hierarchy. Similarly it provides the final user linguistic constructs
to build instances (examples) of structures and to teach the
elaborator of Coq how to take advantage of this knowledge during type
inference. Finally \HB{} lets the library designer improve the
usability of his library by providing alternative interfaces to the
primitive ones, a feature that can also be used to accommodate changes
to the hierarchy without breaking user code. The details of \HB{},
including its compilation to a variant of the Packed Classes
discipline and its implementation in the Elpi extension language for
Coq, are given in~\cite{cohen:hal-02478907}.

In this talk we focus on how to use \HB{} in a Coq development without
any prior knowledge, walking slowly through the example below,
describing the bricks of the language from mixins -- i.e. bare bone building
block of the hierarchy, it packs operations and axioms -- to structures.

\HB{} is available for download at \url{https://github.com/math-comp/hierarchy-builder}.

\paragraph{Example: A hierarchy with a diamond in \HB{}}

We first introduce the base structure of our example,
\verb-CMonoid-, which is defined in two phases: first its contents
using \verb-HB.mixin- and then the structure itself is declared using
\verb-HB.structure.-

\begin{coqcode}
HB.mixin Record CMonoid_of_Type A := { (* The set of axioms making A a commutative monoid. *)
  zero  : A; add   : A -> A -> A;
  addrA : associative add;  (* `add` is associative  *)
  addrC : commutative add;  (* `add` is commutative  *)
  add0r : left_id zero add; (* `zero` is a neutral element *)
}.
HB.structure Definition CMonoid := { A of CMonoid_of_Type A }. (* The structure thereof. *)
Notation "0" := zero.
Infix    "+" := add.

(* The type of the operations and axioms depend on a CMonoid.type structure. *)
Check addrC. (* ?M : CMonoid.type |- commutative (@add ?M) *)
\end{coqcode}

We extend this hierarchy with abelian group, semi ring and ring
structures, which happen to form a diamond (cf
Figure~\ref{fig:hierarchy}), using only the bare minimum information:
we use the same pattern as before to create the abelian group and
semi-ring structures, then a ring being no more than the combination
of the former structures, we do not even need to provide a mixin.

\begin{wrapfigure}[0]{r}{.35\textwidth}
  \vspace{-27em}
  \includegraphics[width=.35\textwidth]{diagram.pdf}
  \vspace{-1.5em}
  \caption{The hierarchy}
  \label{fig:hierarchy}
\end{wrapfigure}

\begin{coqcode}
HB.mixin Record AbelianGrp_of_CMonoid A of CMonoid A := {
  opp   : A -> A;
  (* We can write `add` here since A is a  CMonoid   *)
  addNr : left_inverse zero opp add; (* `opp` is the additive inverse *)
}.
HB.structure Definition AbelianGrp := { A of AbelianGrp_of_CMonoid A }.
Notation "- x"   := (opp x).
Notation "x - y" := (add x (opp y)).

HB.mixin Record SemiRing_of_CMonoid A of CMonoid A := {
  one    : A;
  mul    : A -> A -> A;
  mulrA  : associative mul;  (* `mul` is associative   *)
  mul1r  : left_id one mul;  (* `one` is left neutral  *)
  mulr1  : right_id one mul; (* `one` is right neutral *)
  mulrDl : left_distributive mul add;  (* `mul` distributes over *)
  mulrDr : right_distributive mul add; (*   `add` on both sides  *)
  mul0r  : left_zero zero mul;  (* `zero` is absorbing `mul`     *)
  mulr0  : right_zero zero mul; (*   on both sides               *)
}.
HB.structure Definition SemiRing := { A of SemiRing_of_CMonoid A }.
Notation "1"  := one.
Infix    "*"  := mul.

(* We need no additional mixin to declare the Ring structure. *)
HB.structure Definition Ring := { A of SemiRing_of_CMonoid A & AbelianGrp_of_CMonoid A }.
\end{coqcode}

Remark we can write abstract statements combining successive signatures:
\begin{coqcode}
(* An example statement in the signature of an Abelian group G, combining 0 and -. *)
Check forall G : AbelianGrp.type, forall x : G, x - x = 0.
(* An example statement in the signature of a Semiring S, combining 0, +, and *.  *)
Check forall S : SemiRing.type, forall x : S, x * 1 + 0 = x.
(* An example statement in the signature of a Ring R, combining -, 1 and *.  *)
Check forall R : Ring.type, forall x y : R, x * - (1 * y) = - x * y.
\end{coqcode}

Now, we can equip $\mathbb{Z}$ with commutative monoid, abelian group and
semi ring instances, and \HB{} equips it automatically with the ring
one, because there is only one possible choice.

\begin{coqcode}
HB.instance Definition Z_CMonoid    := CMonoid_of_Type.Build Z 0%Z Z.add
  Z.add_assoc Z.add_comm Z.add_0_l.
HB.instance Definition Z_AbelianGrp := AbelianGrp_of_CMonoid.Build Z Z.opp Z.add_opp_diag_l.
HB.instance Definition Z_SemiRing   := SemiRing_of_CMonoid.Build Z 1%Z Z.mul
  Z.mul_assoc Z.mul_1_l Z.mul_1_r Z.mul_add_distr_r Z.mul_add_distr_l Z.mul_0_l Z.mul_0_r.

(* An example statement in the signature of the Z ring, combining Z, 0, +, -, 1 and * *)
Check forall x : Z, x * - (1 + x) = 0 + 1.
\end{coqcode}

For space constraints we omit the declaration of the dashed arrows in
Figure~\ref{fig:hierarchy}, which are Factories, i.e. alternative
ways of building the same structures, but from different inputs.

Factories are used to accommodate the insertion of intermediate
structures in a hierarchy without breaking user code, as well as to
implement shortcuts: e.g. the factory to build a \coq{Ring} from a
\coq{Type} (rightmost dashed arrow) can omit the commutativity axiom
of the addition, since it is implied by the other axioms.

\bibliography{biblio}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "easychair.tex"
%%% TeX-command-extra-options: "-shell-escape"
%%% ispell-local-dictionary: "english"
%%% End: